MULTI EPOCH SPATIALLY RESOLVED RADIO OBSERVATIONS OF BETELGEUSE'S WIND ACCELERATION REGION

### 2014-06-15 draft1/ms.tex
June 15

\begin{abstract}

Previous VLA observations have only spatially resolved the stellar atmosphere at 0.7\,cm but here we fully resolve the star at all wavelengths between 0.7 and 6.1\,cm. New multi-epoch e-MERLIN data are urgently required.

@@

\cite{lim_1998} used the VLA in its most extended (i.e., A) configuration to resolve Betelgeuse's atmosphere at 0.7\,cm and partially resolve it at 1.3, 2.0, 3.6, and 6\,cm. Because the radio emission is thermal and optically thick, they were able to calculate the mean gas temperature as a function of radius, and found that the temperature decreased steadily from $\sim$3450\,K at 2\,R$_{\star}$, to $\sim$1370\,K at 7\,R$_{\star}$. They also detected an asymmetry in their 0.7\,cm image which they attributed to the action of a large convective cell. To reconcile their results with the extended ultraviolet emission, they concluded that the inner atmosphere must be inhomogeneous to accommodate the hot chromospheric plasma, but that the cooler gas must be 3 orders of magnitude more abundant. \cite{harper_2006} used observations of the chromospheric tracer C II] $\lambda 2325\,\AA$ to confirm this low filling factor for the chromospheric gas.

 lim, richards) (Teff=3600, ohnaka)

Why this paper, i.e. looking for evidence of emerlin
e difference between the peak flux
densities of these two "spots" is about 23 sigmaa

\section{OBSERVATIONS AND DATA REDUCTION}

@@

\label{tab:tab1}

@@

The data were imaged within the Common Astronomical Software Application  \cite[CASA;][]{2007ASPC..376..127M} package.\\
The xxx uvmodelfit was used while bounding the axis ratio between 0 and 1, and the position angle between 0 and 180. \\
Phase center was found using Reid \& Menton\\

\section{RESULTS} 
\subsection{Visibilities} 
\subsection{Radio Images} 
All are axially symmetric
discuss fast switching (carilli paper)
\subsection{Wind Thermal Profile}
give formula

>>>>

\begin{abstract}
We present multi-epoch spatially resolved radio continuum observations of Betelgeuse at various combinations of wavelengths between $0.7-20.5$\,cm.  We used the Very Large Array (VLA) in the A configuration with the Pie Town Very Long Baseline Array (VLBA) antenna to fully resolved its atmosphere at 0.7, 1.3, 2.0, and 6.0\,cm at most epochs.

 New multi-epoch e-MERLIN data are urgently required. Betelgeuse's inner atmosphere has undergone a major morphological change in the past 10 years.

@@

\cite{lim_1998} used the VLA in its most extended (i.e., A) configuration to resolve Betelgeuse's atmosphere at 0.7\,cm and partially resolve it at 1.3, 2.0, 3.6, and 6\,cm. Because the radio emission is thermal and optically thick, they were able to calculate the mean gas temperature as a function of radius. They discovered that the mean gas temperature never reached chromospheric values (i.e., $\gtrsim 5000\,$K) and decreased steadily from $\sim$3450\,K at 2\,R$_{\star}$, to $\sim$1370\,K at 7\,R$_{\star}$. They also detected an asymmetry in their 0.7\,cm image which they attributed to the action of a large convective cell. To reconcile their results with the extended ultraviolet observations, which probed the same regions, they concluded that the inner atmosphere must be inhomogeneous to accommodate the hot chromospheric plasma, but that the cooler gas must be 3 orders of magnitude more abundant. \cite{harper_2006} used observations of the chromospheric tracer C II] $\lambda 2325\,$\AA to confirm this low filling factor for the chromospheric gas.

%have a 24$\sigma$ difference between their peak flux densities,
Recently, an unexpected new discovery by \cite{richards_2013} with e-MERLIN, has revealed that two unresolved radio features dominate the total flux density at 5.2\,cm. The two ``radio hotspots'' are separated by $90\pm 10$\,mas (i.e., $\sim 4\,$R$_{\star}$) and have gas temperatures $T_{\rm{e}} \geq 3800\pm 500\,K$ and $T_{\rm{e}} \geq 5400\pm 600\,K$, with the later value significantly above the photospheric temperature ($T_{\rm{eff}}=3690\pm 54\,$K, \citealt{ohnaka_2011}). The features are so extended that they presumably have either momentum or support and are probably connected with the mass-loss mechanism in Betelgeuse. The potential importance of these radio hotspots has been the motivation for this paper. In the following sections we present, analyze, and discuss multi-epoch, mutli-wavelength radio observations of Betelgeuse, taken $\sim 10$ years prior to the e-MERLIN observations. At short wavelengths, our data has comparable or superior resolution to the e- MERLIN data and so the goal of the paper is to search for signatures of these features to improve our understanding of them.

\section{OBSERVATIONS AND DATA REDUCTION}
Betelgeuse was observed with the NRAO\footnote{The National Radio Astronomy Observatory is a facility of
the National Science Foundation operated under cooperative agreement by Associated Universities, Inc.} VLA  in the A configuration with the Pie Town (PT) Very Long Baseline Array (VLBA) antenna during the period December 2000 and October 2004. The inclusion of the PT antenna to the VLA A configuration improves the east-west resolution by a factor of 2. Our observations are summarized in the first two columns of Table \ref{tab1}. Our two multi-wavelength data sets in 2002 (Program: AH0778) were taken only $\sim 2$ months apart and consist of all wavelength bands between 1.3 and 20.5\,cm. We observed the star again under program AH0824 in 2003 and 2004 at all wavelength bands between 0.7 and 20.5\,cm. We also obtained VLA A configuration plus PT antenna data from the NRAO data archive (Program: AL0525) at  0.7 and 1.3\,cm which were taken in late 2000 and early 2001, respectively.

The data were imaged within the Common Astronomical Software Application  \cite[CASA;][]{2007ASPC..376..127M} package. calibrators, fast switching, cycle times

\subsection{Visibilities}

The xxx uvmodelfit was used while bounding the axis ratio between 0 and 1, and the position angle between 0 and 180. \\
Phase center was found using Reid \& Menton\\

\subsection{Imaging}
Quote Tables synthesized beam and rms'

@@

\label{tab1}

@@

\section{RESULTS} 
\subsection{Radio Maps} 
All are axially symmetric
threw away Q band pie town baselines
\subsection{Diameters} 
Previous VLA observations have only spatially resolved the stellar atmosphere at 0.7\,cm but here we fully resolve the star at all wavelengths between 0.7 and 6.1\,cm.
\subsection{Flux Densities}
plot flux versus frequency
\subsection{Temperature Profiles}
give formula
